Level-Set method on \texttt{p4est}

### 2015-06-19 paper/tex/applications.tex
Last edits for this time

The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good synthesis of the performance of the various algorithms. We monitored the performance of the code over five time iterations, as presented in Algorithm \ref{alg:stefan}, for two different maximum resolutions. In both cases, the forest is built on a $20\times20\times20$ macro-mesh. The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points. The results are presented in figure \ref{fig:stefan_scaling}. As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (c.f.\ table \ref{tab:scaling_stefan}).

>>>>

The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good synthesis of the performance of the various algorithms. We monitored the performance of the code over five time iterations, as presented in Algorithm \ref{alg:stefan}, for two different maximum resolutions. In both cases, the forest is built on a $20\times20\times20$ macro-mesh. The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points. The results are presented in figure \ref{fig:stefan_scaling}. As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (cf.\ Table \ref{tab:scaling_stefan}).

### 2015-06-19 paper/tex/conclusion.tex
Last edits for this time

In particular we showed that the scalability of the semi-Lagrangian algorithm, depends on the CFL number. Great scalability is observed for intermediate CFL numbers, e.g. $\text{CFL}\sim10$. At higher CFL numbers, however, the departure points are potentially further dispersed across processors, which limits the scalability. This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account. A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g. by using a forward-in-time integration of grid points. Such an estimate could also reduce the number of semi-Lagrangian iterations. These ideas are postponed for further investigations. We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation. Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors.

>>>>

In particular we showed that the scalability of the semi-Lagrangian algorithm, depends on the CFL number. Great scalability is observed for intermediate CFL numbers, e.g. $\text{CFL}\sim10$. At higher CFL numbers, however, the departure points are potentially further dispersed across processors, which limits the scalability. This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account. A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g.\ by using a forward-in-time integration of grid points. Such an estimate could also reduce the number of semi-Lagrangian iterations. These ideas are postponed for further investigations. We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation. Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors.

### 2015-06-19 paper/tex/scaling_results.tex
Last edits for this time

\subsection{Interpolation}
In this section we show the results for a simple test to measure the scalability of the interpolation Algorithm \ref{alg:interpolation}. The test consists of interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions. We consider two cases, a small test on a level\footnote{the level is the number of recursive splits allowed for each tree.} 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes. In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stabilized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.

To simulate the effect of different CFL numbers, we generate the random points such that on each process $\alpha$ percentage of them are located outside the process boundary and thus will initiate communication. Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in figure \ref{fig:interpolation}. We also present a third row of results for a much larger problem with roughly 1.66B nodes on a level 14 tree. Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote process. For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $2048$ processes when $\alpha = 95\%$ (c.f\ table \ref{tab:scaling_interpolation}). Note, however, that this is expected since the total time is dominated by communication for $\alpha = 95\%$ and there is very little local work in this case. Indeed, the last row of figure \ref{fig:interpolation} show much better scaling behavior on a larger problem size, e.g. efficiencies are increased from $e=34\%$ to $e=68\%$ for $\alpha = 95\%$ on 4096 processes. This is a typical result with strong scaling and simply implies that our algorithms are scalablable for sufficiently large problems.
\begin{figure}[htbp]
	\begin{center}
		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_small_alpha_5.pdf}} \hspace{0.5 in}
		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_small_alpha_95.pdf}}
		\\
		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_large_alpha_5.pdf}} \hspace{0.5 in}
		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_large_alpha_95.pdf}}
		\\
		\subfigure[$N_G = 1.66$B, $\alpha = 50\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_super_large_alpha_50.pdf}} \hspace{0.5 in}
		\subfigure[$N_G = 1.66$B, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_super_large_alpha_95.pdf}}
	\end{center}
	\caption{Strong scaling of Algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each process. Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (c.f. Algorithm \ref{alg:interpolation}). The black dashed line represents the ideal scaling. The results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c). For the extreme case (d) the algorithm stops scaling at $2048$ processes due to communication overhead. Note, however, that this merely indicates that the problem size is not large enough for this test case. Indeed much better scaling is obtained when the problem size is increased to $N_G = 1.66$B points (e-f).}
	\label{fig:interpolation}
\end{figure}

@@

Table \ref{tab:semilagrangian} illustrates the dependence of the number of sub-iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance, which necessitates more sub-iterations to generate the grid. Note that the apparent dependence on the number of processes is most likely due to round-off errors during the interpolation and/or backtracking steps. Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively. To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of sub-iterations required for the grid construction as reported in table \ref{tab:semilagrangian}. For both problems, excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$. The algorithm even shows good scalability when taken to the extreme, i.e. for $\text{CFL} = 100$.

An increase in the CFL number has two effects on the algorithm. First, a larger fraction of the departure points lands in the domains of remote processes. Moreover, these points are potentially dispersed across a larger number of processes. This means that the communication volume should increase with the CFL number. Second, as more points are shipped to remote processes for interpolation, there is a greater chance that the interpolation load is imbalanced across processes. This is especially true for regions of space in which the streamlines cluster. Both factors can contribute to reducing the scalability of the algorithm at large CFL numbers. 
\begin{figure}[htbp]
	\begin{center}
		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_1.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_10.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_100.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
	\end{center}
	\caption{Strong scaling of a single time step of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes. Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers. The solid dashed line represents the ideal scaling. Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (c.f. table \ref{tab:semilagrangian}). Here \texttt{p4est\_nodes\_new}, \texttt{p4est\_partition} and \texttt{p4est\_refine\_coarsen} refer to constructing the global indexing for nodes, partitioning the forest, and the refining/coarsening operation, respectively \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.}
	\label{fig:semilagrangian_small}
\end{figure}

\begin{figure}[htbp]
	\begin{center}
		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_1.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_10.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_100.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_1.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_10.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_100.pdf}}
	\end{center}
	\caption{Strong scaling of a single time step of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes. Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers. The solid dashed line represents the ideal scaling. Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (c.f. table \ref{tab:semilagrangian}).}
	\label{fig:semilagrangian_large}
\end{figure}

@@

To better understand the importance of the CFL number on the scalability, we have recorded a complete history of the communication pattern in the interpolation step. Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processes}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in megabytes (MB), for $p=4096$ processes. Furthermore, these values are reported for the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm. There are several points to make. First, increasing the CFL number greatly increases the load imbalance, as shown by the spread of the data in figure \ref{fig:communication_4096_pf}. This is because at higher CFL numbers, it is more likely that some processes will receive a larger portion of the backtracked points. Second, increasing the CFL number increases both the communication volume and its spread across processes (c.f. figure \ref{fig:communication_4096_vf}). Interestingly, however, the number of sent and received messages do not seem to affected by the CFL number. The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last sub-iteration of the semi-Lagrangian algorithm. This can be justified by noting that for the final sub-iteration, the partitioning of $G^{n+1}$ is more consistent with the partitioning of the departure points on $G^n$. Detailed information about the load balancing and the communication patterns is listed in table \ref{tab:communication}.
\begin{figure}[htbp]
	\begin{center}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_point_first_4096.pdf}  \label{fig:communication_4096_pf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_message_first_4096.pdf}\label{fig:communication_4096_mf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_volume_first_4096.pdf} \label{fig:communication_4096_vf}}
		\\
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_point_last_4096.pdf}   \label{fig:communication_4096_pl}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_message_last_4096.pdf} \label{fig:communication_4096_ml}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_volume_last_4096.pdf}  \label{fig:communication_4096_vl}}		
	\end{center}
	\caption{Performance indicators of the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm for the level-set advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}). Increasing the CFL number greatly causes load imbalance during interpolation (a) and increases the communication volume (c). However, the CFL number does not seem to appreciably affect the number of messages sent by the processes (b). During the last semi-Lagrangian sub-iteration, the initial grid $G_0$ (c.f. Algorithm \ref{alg:semi-lagrangian}) is very close to the final grid. As a result, the load imbalance is considerably improved (d). Curiously, however, the communication pattern does not seem to be change much between first and last sub-iterations (e,f).}
	\label{fig:communication_4096}
\end{figure}

@@

Finally we present the scaling results of our parallel reinitialization algorithm where we extensively make use of Algorithm \ref{alg:overlap} for overlapping the computations with the communications when computing spatial derivatives. Our test consists in computing the signed distance function to a collection of 100 spheres, whose radii and centers are chosen randomly. The test is performed on a small, level-8 Octree with about 21M and a larger, level-10 Octree with about 337M grid points. In both cases the forest is built on a $3\times3\times3$ macro-mesh. Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in Algorithm \ref{alg:overlap}, scales very well (c.f.\ table \ref{tab:scaling_reinit}). In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize Algorithm \ref{alg:overlap}. 
\begin{figure}
\centering
\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Reinit_Small.pdf}}
\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Reinit_Large.pdf}}
\caption{Scalability of the reinitialization test for a small (left) and large (right) Octree with roughly 21M and 337M grid points, respectively. The black dashed line represents ideal scaling. Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (c.f. algorithm \ref{alg:overlap}).}
\label{fig:reinit}
\end{figure}

>>>>

\subsection{Interpolation}
In this section we show the results for a simple test to measure the scalability of the interpolation Algorithm \ref{alg:interpolation}. The test consists of interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions. We consider two cases, a small test on a level\footnote{The level is the number of recursive splits allowed for each tree.} 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes. In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stabilized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.

To simulate the effect of different CFL numbers, we generate the random points such that on each process $\alpha$ percentage of them are located outside the process boundary and thus will initiate communication. Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in Figure \ref{fig:interpolation}. We also present a third row of results for a much larger problem with roughly 1.66B nodes on a level 14 tree. Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote process. For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $2048$ processes when $\alpha = 95\%$ (cf.\ Table \ref{tab:scaling_interpolation}). Note, however, that this is expected since the total time is dominated by communication for $\alpha = 95\%$ and there is very little local work in this case. Indeed, the last row of Figure \ref{fig:interpolation} show much better scaling behavior on a larger problem size, e.g. efficiencies are increased from $e=34\%$ to $e=68\%$ for $\alpha = 95\%$ on 4096 processes. This is a typical result with strong scaling and simply implies that our algorithms are scalablable for sufficiently large problems.
\begin{figure}[htbp]
	\begin{center}
		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_small_alpha_5.pdf}} \hspace{0.5 in}
		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_small_alpha_95.pdf}}
		\\
		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_large_alpha_5.pdf}} \hspace{0.5 in}
		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_large_alpha_95.pdf}}
		\\
		\subfigure[$N_G = 1.66$B, $\alpha = 50\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_super_large_alpha_50.pdf}} \hspace{0.5 in}
		\subfigure[$N_G = 1.66$B, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_super_large_alpha_95.pdf}}
	\end{center}
	\caption{Strong scaling of Algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each process. Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (cf.\ Algorithm \ref{alg:interpolation}). The black dashed line represents the ideal scaling. The results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c). For the extreme case (d) the algorithm stops scaling at $2048$ processes due to communication overhead. Note, however, that this merely indicates that the problem size is not large enough for this test case. Indeed much better scaling is obtained when the problem size is increased to $N_G = 1.66$B points (e-f).}
	\label{fig:interpolation}
\end{figure}

@@

Table \ref{tab:semilagrangian} illustrates the dependence of the number of sub-iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance, which necessitates more sub-iterations to generate the grid. Note that the apparent dependence on the number of processes is most likely due to round-off errors during the interpolation and/or backtracking steps. Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively. To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of sub-iterations required for the grid construction as reported in Table \ref{tab:semilagrangian}. For both problems, excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$. The algorithm even shows good scalability when taken to the extreme, i.e. for $\text{CFL} = 100$.

An increase in the CFL number has two effects on the algorithm. First, a larger fraction of the departure points lands in the domains of remote processes. Moreover, these points are potentially dispersed across a larger number of processes. This means that the communication volume should increase with the CFL number. Second, as more points are shipped to remote processes for interpolation, there is a greater chance that the interpolation load is imbalanced across processes. This is especially true for regions of space in which the streamlines cluster. Both factors can contribute to reducing the scalability of the algorithm at large CFL numbers. 
\begin{figure}[htbp]
	\begin{center}
		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_1.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_10.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_100.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
	\end{center}
	\caption{Strong scaling of a single time step of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes. Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers. The solid dashed line represents the ideal scaling. Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (cf.\ Table \ref{tab:semilagrangian}). Here \texttt{p4est\_nodes\_new}, \texttt{p4est\_partition} and \texttt{p4est\_refine\_coarsen} refer to constructing the global indexing for nodes, partitioning the forest, and the refining/coarsening operation, respectively \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.}
	\label{fig:semilagrangian_small}
\end{figure}

\begin{figure}[htbp]
	\begin{center}
		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_1.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_10.pdf}}
		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_100.pdf}}
		\\
		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_1.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_10.pdf}}
		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_100.pdf}}
	\end{center}
	\caption{Strong scaling of a single time step of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes. Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$. Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers. The solid dashed line represents the ideal scaling. Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (cf.\ Table \ref{tab:semilagrangian}).}
	\label{fig:semilagrangian_large}
\end{figure}

@@

To better understand the importance of the CFL number on the scalability, we have recorded a complete history of the communication pattern in the interpolation step. Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processes.}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in megabytes (MB), for $p=4096$ processes. Furthermore, these values are reported for the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm. There are several points to make. First, increasing the CFL number greatly increases the load imbalance, as shown by the spread of the data in Figure \ref{fig:communication_4096_pf}. This is because at higher CFL numbers, it is more likely that some processes will receive a larger portion of the backtracked points. Second, increasing the CFL number increases both the communication volume and its spread across processes (cf.\ Figure \ref{fig:communication_4096_vf}). Interestingly, however, the number of sent and received messages do not seem to be affected by the CFL number. The bottom row of Figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last sub-iteration of the semi-Lagrangian algorithm. This can be justified by noting that for the final sub-iteration, the partitioning of $G^{n+1}$ is more consistent with the partitioning of the departure points on $G^n$. Detailed information about the load balancing and the communication patterns is listed in Table \ref{tab:communication}.
\begin{figure}[htbp]
	\begin{center}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_point_first_4096.pdf}  \label{fig:communication_4096_pf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_message_first_4096.pdf}\label{fig:communication_4096_mf}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_volume_first_4096.pdf} \label{fig:communication_4096_vf}}
		\\
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_point_last_4096.pdf}   \label{fig:communication_4096_pl}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_message_last_4096.pdf} \label{fig:communication_4096_ml}}
		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_volume_last_4096.pdf}  \label{fig:communication_4096_vl}}		
	\end{center}
	\caption{Performance indicators of the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm for the level-set advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}). Increasing the CFL number causes load imbalance during interpolation (a) and increases the communication volume (c). However, the CFL number does not seem to appreciably affect the number of messages sent by the processes (b). During the last semi-Lagrangian sub-iteration, the initial grid $G_0$ (cf.\ Algorithm \ref{alg:semi-lagrangian}) is very close to the final grid. As a result, the load imbalance is considerably improved (d). Curiously, however, the communication pattern does not seem to be change much between first and last sub-iterations (e,f).}
	\label{fig:communication_4096}
\end{figure}

@@

Finally we present the scaling results of our parallel reinitialization algorithm where we extensively make use of Algorithm \ref{alg:overlap} for overlapping the computations with the communications when computing spatial derivatives. Our test consists in computing the signed distance function to a collection of 100 spheres, whose radii and centers are chosen randomly. The test is performed on a small, level-8 Octree with about 21M and a larger, level-10 Octree with about 337M grid points. In both cases the forest is built on a $3\times3\times3$ macro-mesh. Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in Algorithm \ref{alg:overlap}, scales very well (cf.\ Table \ref{tab:scaling_reinit}). In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize Algorithm \ref{alg:overlap}. 
\begin{figure}
\centering
\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Reinit_Small.pdf}}
\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Reinit_Large.pdf}}
\caption{Scalability of the reinitialization test for a small (left) and large (right) Octree with roughly 21M and 337M grid points, respectively. The black dashed line represents ideal scaling. Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (cf.\ Algorithm \ref{alg:overlap}).}
\label{fig:reinit}
\end{figure}
